---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr:::opts_chunk$set(echo = TRUE)
```

```{r}
#1
upperCase <- LETTERS
lowerCase <- letters
print(upperCase)
print(lowerCase)
```
#upperCase
# [1] "A" "B" "C" "D" "E" "F" "G" "H" "I" "J" "K" "L" "M" "N" "O" "P" "Q" "R" "S" "T" "U" "V" "W"
#[24] "X" "Y" "Z"
#lowerCase
# [1] "a" "b" "c" "d" "e" "f" "g" "h" "i" "j" "k" "l" "m" "n" "o" "p" "q" "r" "s" "t" "u" "v" "w"
#[24] "x" "y" "z"

#1a.
```{r}
x <- head(LETTERS, 11)
print(x)
```
# [1] "A" "B" "C" "D" "E" "F" "G" "H" "I" "J" "K"

#1b.
```{r}
odd_letters <- seq(from = 1, to = length(LETTERS),by = 2)
odd <- LETTERS[odd_letters]
print(odd)
```
# [1] "A" "C" "E" "G" "I" "K" "M" "O" "Q" "S" "U" "W" "Y"

#1c.
```{r}
vowels <- c("A", "E", "I", "O", "U")
print(vowels)
```
# [1] "A" "E" "I" "O" "U"

#1d.
```{r}
y <- tail(letters, 5)
print(y)
```
#[1] "v" "w" "x" "y" "z"

#1e.
```{r}
letter <- letters[15:24]
print(letter)
```
# [1] "o" "p" "q" "r" "s" "t" "u" "v" "w" "x"

#2.
```{r}
temperature <- c(avg_temp_tuguegarao = 42, avg_temp_manila = 39, avg_temp_iloilo = 34, avg_temp_tacloban = 34, avg_temp_samal_island = 30, avg_temp_davao_city = 27)
print(temperature)
```
#avg_temp_tuguegarao       avg_temp_manila       avg_temp_iloilo     avg_temp_tacloban
#                   42                    39                    34                    34
#avg_temp_samal_island   avg_temp_davao_city
#                   30                    27

#2a.
```{r}
city <- c("Tuguegarao City","Manila","Iloilo City","Tacloban","Samal Island","Davao City")
print(city)
```
#[1] "Tuguegarao City" "Manila"          "Iloilo City"     "Tacloban"        "Samal Island"  
#[6] "Davao City"

#2b.
```{r}
temp <- c(42, 39, 34, 34, 30, 27)
print(temp)
```
#[1] 42 39 34 34 30 27

#2c.
```{r}
df <- data.frame(city,temp)
print(df)
```
#            city   temp
#1 Tuguegarao City   42
#2          Manila   39
#3     Iloilo City   34
#4        Tacloban   34
#5    Samal Island   30
#6      Davao City   27

#2d.
```{r}
names(df) <- c("City", "Temperature")
print(df)
```
#             City Temperature
#1 Tuguegarao City          42
#2          Manila          39
#3     Iloilo City          34
#4        Tacloban          34
#5    Samal Island          30
#6      Davao City          27

#2e.
```{r}
str(df)
```
#'data.frame': 6 obs. of  2 variables:
# $ City       : chr  "Tuguegarao City" "Manila" "Iloilo City" "Tacloban" ...
# $ Temperature: num  42 39 34 34 30 27
# The data frame has 6 obs and 2 variables.The data type of the variable "City" has a factor with 6 levels and the "Temperature" is a numeric variable.

#2f.
```{r}
df[3,]
```
#for row 3
#         City Temperature
# 3 Iloilo City          34
df[4,]
#      City Temperature
# 4 Tacloban          34

#2g.
```{r}
df[df$Temperature==max(df$Temperature), "City"]
df[df$Temperature==min(df$Temperature), "City"]
```
#[1] "Tuguegarao City"
#[1] "Davao City"


USING MATRICES

#2a.
```{r}
mat <- matrix(c(1:8, 11:14), nrow=3, ncol=4, byrow = TRUE)
mat
```
#      [,1] [,2] [,3] [,4]
# [1,]    1    2    3    4
# [2,]    5    6    7    8
# [3,]   11   12   13   14

#2b.
```{r}
mat <- matrix(c(1:8, 11:14), nrow=3, ncol=4, byrow = TRUE)
mat_multiply <- mat*2
mat_multiply
```
#      [,1] [,2] [,3] [,4]
# [1,]    2    4    6    8
# [2,]   10   12   14   16
# [3,]   22   24   26   28

#2c.
```{r}
mat[2,]
```
# [1] 5 6 7 8

#2d.
```{r}
mat[1:2, 3:4]
```
#      [,1] [,2]
# [1,]    3    4
# [2,]    7    8

#2e.
```{r}
mat[3,2:3]
```
# [1] 12 13

#2f.
```{r}
mat[,4]
```
# [1]  4  8 14

#2g.
```{r}
rownames(mat_multiply) <- c("isa", "dalawa", "tatlo")
colnames(mat_multiply) <- c("uno", "dos", "tres", "quatro")
mat_multiply
```
#        uno dos tres quatro
# isa      2   4    6      8
# dalawa  10  12   14     16
# tatlo   22  24   26     28

#2h.
```{r}
dim(mat) <- c(6,2)
mat
```
#      [,1] [,2]
# [1,]    1    3
# [2,]    5    7
# [3,]   11   13
# [4,]    2    4
# [5,]    6    8
# [6,]   12   14


USING ARRAYS

#3a.
```{r}
p <- rep(c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1), each=2)
a <- array(p,dim=c(2,4,3))
a
```
# , , 1

#     [,1] [,2] [,3] [,4]
# [1,]    1    2    3    6
# [2,]    1    2    3    6

# , , 2

#     [,1] [,2] [,3] [,4]
# [1,]    7    8    9    0
# [2,]    7    8    9    0

# , , 3

#     [,1] [,2] [,3] [,4]
# [1,]    3    4    5    1
# [2,]    3    4    5    1

#3b.
```{r}
Array has 3 dimensions
```

#3c.
```{r}
p <- rep(c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1), each=2)
h <- array(p,dim=c(2,4,3))
row_names <- letters[1:2]
col_names <- LETTERS[1:4]
dimnames(h) <- list(row_names, col_names, c("1st-Dimensional Array", "2nd-Dimensional Array", "3rd-Dimensional Array"))
h
```
#, , 1st-Dimensional Array

#  A B C D
# a 1 2 3 6
# b 1 2 3 6

# , , 2nd-Dimensional Array

#  A B C D
# a 7 8 9 0
# b 7 8 9 0

# , , 3rd-Dimensional Array

#   A B C D
# a 3 4 5 1
# b 3 4 5 1